---
title: "Introduction to DIMESS using cancer scRNA data"
author: "Peter Boennighausen"
output:
  html_document:
    theme: united
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(
    cache = TRUE,
    cache.lazy = FALSE,
    tidy = TRUE, 
    warning = FALSE,
    message = FALSE
)
```
DIMESS is a package for analyzing scRNA data. It allows you to try out different
distance metrics easily on Seurat objects in order to cluster genes and explore gene
expression networks. 

We start by loading some packages. The most important is Seurat -- DIMESS expects to interact with Seurat objects
throughout the entire workflow.
```{r load packages}
library(Seurat)
library(readxl)
library(dplyr)
source("dimess.R")
```

Now that we have the necessary packages, we load data. Some of this is example specific, such as the code to attach
metadata to the Seurat object. When working with your own datasets, be sure to ensure that your Seurat object has count data 
with the proper feature (gene) names.

The normalization procedure is basically identical to that described in the Seurat vignettes -- the point of 
DIMESS is to augment rather than replace your existing Seurat workflow.
```{r load data and metadata, results='hide'}
load("./adenoallraw.RData")
type <- "tumor"
lung <- CreateSeuratObject(adeno.all$TUMOR, min.cells = 3, min.features = 0, project = "nsclc")

metadata <- read_excel("./MetaData.xlsx", sheet = 1)
strsplit.ind2 <- seq(from=2, by=2, length.out = nrow(metadata))
metadata <- metadata %>%
    mutate(piece = unlist(strsplit(Patient_piece, split="-"))[strsplit.ind2]) %>%
    filter(CellType == type)
lung <- AddMetaData(object = lung, metadata = metadata)
lung <- NormalizeData(object = lung, normalization.method = "LogNormalize", scale.factor = 1e4)
all.genes <- rownames(x = lung)
lung <- ScaleData(object = lung, features = all.genes)
```
Now that we have a Seurat object, let's examine gene expression networks. First, we
identify a subset of cell _names_ that interest us. In our case, we're looking
at only patients 3 and 4, so we get the cell names of interest by subsetting
the metadata. 
```{r subset cells and build gene graphs, results='hide'}
cells <- metadata[metadata$'PatientNumber MS' %in% c(3:4),]$cell
```
Next, we identify some list of metrics that interest up. DIMESS supports the 17 
from the `dismay` package, here we use almost all of them.
```{r list metrics}
metrics <- c('pearson', 'spearman', 'kendall', 'bicor', 'zi_kendall', 'binomial', 'MI',
  'cosine', 'jaccard', 'canberra', 'euclidean', 'manhattan',
  'weighted_rank', 'hamming')
```
We then pass our `cells` and `metrics` to DIMESS's `build_gene_graphs` function, 
which returns a list of `tbl_graph` items that
can be passed to the other functions in the package. 
```{r build graphs}
graphs <- build_gene_graphs(lung, cells, metrics)
```

For example, we can plot some of the clusters for different metrics.
```{r plot all clusters}
for (metric in c("pearson", "MI", "bicor")) {
    plot_gene_clusters(graphs[[metric]], metric)
}
```

We could also plot the overall network, colored by centrality:
```{r plot-network}
for (metric in c("cosine", "jaccard", "manhattan")) {
    plot_network(graphs[[metric]], metric)
}
```

We can also find just the most central genes:
```{r top-10}
for (metric in c("euclidean", "binomial", "weighted_rank")) {
  print(most_central(graphs[[metric]], 10))
}
```
